### 2020-12-10 doc/manuscript/Dow_manuscript_Nature.Rmd
start reformatting for Nature
#41
@camerondow35 , I created a new version rather than overwriting the GCB version

---
title: ''
output:
  word_document
  #bookdown::word_document2: 
    #reference_docx: word-styles-reference.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
    #fig_caption: yes
bibliography: references.bib
csl: global-change-biology.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Target journal: Global Change Biology
## Instructions for authors: https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/page/journal/13652486/homepage/forauthors.html
## Primary Research Articles present the results of a completed research project and are up to 8000 words in length. Word limits apply to the main body of the text (Introduction, Materials and Methods, Results, Discussion, and Acknowledgements).

## I don't think GCB has an official limit. I just saw a recently accepted article with 10 figures and 4 tables, so we're fine!

@@

#Table numbers
hypothesis_table=1
table_species=2
mean_growth_parameters_table=3

#Figure numbers
schematic=1 #illustrates parameters
growth_curves_all = 2  
    rel_growth_SCBI = "2a" 
    rel_growth_HF = "2b" 
    cum_growth_SCBI = "2c" 
    cum_growth_HF = "2d" 
climwin = 3
pheno_Tsensitivity = 4
DOYtiming = 5
DOYtiming_SCBI = "5a"
DOYtiming_Harvard = "5b"
tree_rings = 6

@@

growth_curves_all_caption  <- "Figure 2. Growth trajectories for ring- and diffuse-porous trees, as both relative and cumulative fractions of total annual growth. Each line represents one year's growth for a given tree, fit with McMahon model."

@@

tree_rings_caption <- "Figure 6. Climate sensitivity of annual growth, as derived from tree-rings, for dominant ring- and diffuse- porous species at the Smithsonian Conservation Biology Institute and Harvard Forest. Shown are Pearson correlations between ring-width index and and monthly Tmax. Species codes are as given in Table 2."

@@

#SI table captions
sample_size_by_year_caption <- "Table S1. Sample size by year"

@@

# Abstract
```{r eval=FALSE, echo=FALSE}
## For GCB:
## Abstract limit for GCB is 300 words. 
## 6-10 keywords

@@

As the climate changes, warmer spring temperatures are causing earlier leaf-out in temperate deciduous forests, along with earlier initiation of the annual period where these ecosystems are net sinks of carbon dioxide (CO~2~).
As a result, there is a tendency towards increased growing season length and annual CO~2~ sequestration.
However, little is known about how spring temperatures affect tree stem growth, which consumes a modest portion of the carbon fixed annually but is critical in that wood has a long residence time in the ecosystem.
We need to know how the timing and rate of woody growth are affected by warming spring temperatures.
Here, we show that in two temperate deciduous forests warmer maximum temperatures in the spring shift growth milestones earlier in the year, but have only a modest effect on peak growing season length and little effect on annual growth.
This contrasts with the expectation *in most models* that an earlier start to growing season would significantly increase the rate of biomass accumulation and thereby enhance the CO~2~ sink of temperate deciduous forests.
It is, however, consistent with most tree-ring studies in temperate deciduous forests showing that annual growth tends to be more sensitive to climatic conditions during the peak growing season than in the spring, with eddy-covariance records showing that annual CO~2~ sequestration is more strongly linked to conditions in the peak growing season than to growing season length, and with evidence that forest CO~2~ sequestration is sink-limited.
Thus, our findings indicate that warming spring conditions are unlikely to impact temperate deciduous forest growth sufficiently to produce any substantial negative climate change feedback.

**Keywords**:

@@

# Introduction

@@

Reducing this uncertainty requires better understanding of the drivers of tree growth [@bugmann_tree_2019]. Tree growth is sensitive to climatic factors like precipitation, temperature, and cloudiness in early growing season [@Helcoski_growing_2019]; [@Ciais_europe-wide_2005]; [@Goulden_exchange_1996], as well as other physical factors including nitrogen content of soil, increasing atmospheric CO~2~, deposition of atmospheric pollutants in acid rain [@mathias_disentangling_2018], and invasion by pests or pathogens. Understanding how growth is influenced by these factors is essential for developing accurate models to inform forest managers and policy makers about the role of future forests in climate change mitigation efforts.

## (spring temperatures are rising, and this is affecting leaf phenology and carbon cycling)
In recent decades, an increase in average spring temperature has been impacting many natural systems including forests ([@parmesan_globally_2003]; [@zhizong_fingerprinting_2006]). 
The effects of this spring warming are most apparent when examining leaf phenology shifts, *which have been extensively studied ([@way_tree_2011]; [@klosterman_later_2018]), leading to many critical observations*. For example, some regions have seen earlier leaf out in deciduous trees, which is often matched or exceeded by delays in leaf senescence timing, when springtime temperatures are above average [@ibanez_forecasting_2010];@jeong_phenology_2011];@crabbe_extreme_2016]. The emergence of leaves signals the beginning of carbon sequestration or in other words, the growing season of a tree. It is believed that the longer the tree has leaves - or as the growing season is extended - the net annual carbon sequestered will increase as well (*i,e*, net ecosystem exchange, $NEE$; @churkina_spatial_2005; @keenan_net_2014; @zhou_explaining_2016].)

However, more of the interannual variation in forest $NEE$ is explained by variation in maximum C uptake rates than by the length of the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019]. This implies that earlier leafout will lead to smaller increases in NEE than changes in other phenological traits like the maximum C uptake rate. This idea agrees with patterns seen in tree cores, where peak growing season climate was more influential than spring climate [@Helcoski_growing_2019] but many models still predict the extension of growing season will increase NEE. A paper by @zani_increased_2020 found that earlier leaf out led to earlier leaf senescence within the deciduous trees in their study, reducing the ability of the trees to sequester significantly higher levels of C. Instead they posit that senescence is induced when seasonal carbon-sink capacity of the tree is reached, no matter how extended the potential growing season is. This result directly contradicts those previous predictions that warmer Spring temperatures and earlier leaf out will lead to drastically increased C sequestration.

@@

# Materials and Methods

@@

# Results

## Phenological patterns

## Critical temperature windows

## Climate sensitivity of phenology

## Climate sensitivity of growth

# Discussion

Indeed, the tree-ring records reveal that at SCBI the strongest temperature sensitivity aligned with peak growing season [@helcoski_growing_2019], whereas there were fewer significant climate correlations for the species at Harvard Forest (Fig. `r tree_rings`).
Relative to other temperate deciduous forest sites, the pattern at SCBI tends to be more typical: annual growth is most commonly controlled by climate conditions during the peak growing season.
(*1-2 sentences generalizing about climate sensitivity of temperate deciduous forest growth--Neil can do this most easily.*)
Thus, while our analysis is limited to two sites, we expect that the finding that spring temperatures have limited effect on annual tree growth will hold generally.

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (REFS), an earlier start to net C sequestration (REFS), and--as shown here--earlier initiation of stem growth (Figs. `r climwin`- `r DOYtiming`).
Yet, an earlier start to the growing season does not imply a longer growing season [Table `r hypothesis_table`, Figs. `r pheno_Tsensitivity`, `r DOYtiming`;  @zani_increased_2020].
Rather, growth finishes sooner-- presumably because of *C sink limitation* [@zani_increased_2020].

>>>>

---
title: ''
output:
  word_document
  #bookdown::word_document2: 
    #reference_docx: word-styles-reference.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
    #fig_caption: yes
bibliography: references.bib
csl: nature.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Target journal: Nature

## Nature articles: 3-4 modest display items

@@

#Table numbers
hypothesis_table=1
table_species=2  #goes to SI
mean_growth_parameters_table=3

#Figure numbers
schematic=1 #illustrates parameters
growth_curves_all = 2  #goes to SI
    rel_growth_SCBI = "2a" 
    rel_growth_HF = "2b" 
    cum_growth_SCBI = "2c" 
    cum_growth_HF = "2d" 
climwin = 3 # probably stays?
pheno_Tsensitivity = 4 
DOYtiming = 5
DOYtiming_SCBI = "5a"
DOYtiming_Harvard = "5b"
tree_rings = 6  #stays, hopefully adding a lot more sites

@@

growth_curves_all_caption  <- "Figure 2.  Growth trajectories for ring- and diffuse-porous trees, as both relative and cumulative fractions of total annual growth. Each line represents one year's growth for a given tree, fit with McMahon model. (THIS FIGURE GOES TO SI.)"

@@

tree_rings_caption <- "Figure 6. Climate sensitivity of annual growth, as derived from tree-rings, for dominant ring- and diffuse- porous species at the Smithsonian Conservation Biology Institute and Harvard Forest. Shown are Pearson correlations between ring-width index and and monthly Tmax. Species codes are as given in Table 2.(THIS FIGURE STAYS, PROBABLY MODIFIED SIGNFICANTLY TO INCLUDE LOTS MORE SITES.)"

@@

#SI table captions
sample_size_by_year_caption <- "Table S1. Sample size by year (THIS TABLE GOES TO SI)"

@@

# (Summary paragraph)
```{r eval=FALSE, echo=FALSE}
## For Nature:
# Articles start with a fully referenced summary paragraph, ideally of no more than 200 words, which is separate from the main text and avoids numbers, abbreviations, acronyms or measurements unless essential. It is aimed at readers outside the discipline. This summary paragraph should be structured as follows: 2-3 sentences of basic-level introduction to the field; a brief account of the background and rationale of the work; a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent); and finally, 2-3 sentences putting the main findings into general context so it is clear how the results described in the paper have moved the field forwards. Please refer to our annotated example  to see how the summary paragraph should be constructed.

@@

(*word limit 200; this is currently 266*)

As the climate changes, warmer spring temperatures are causing earlier leaf-out in temperate deciduous forests, along with earlier initiation of the annual period where these ecosystems are net sinks of carbon dioxide (CO~2~).
As a result, there is a tendency towards increased growing season length and annual CO~2~ sequestration.
However, little is known about how spring temperatures affect tree stem growth, which consumes a modest portion of the carbon fixed annually but is critical in that wood has a long residence time in the ecosystem.
We need to know how the timing and rate of woody growth are affected by warming spring temperatures.
Here, we show that in two temperate deciduous forests warmer maximum temperatures in the spring shift growth milestones earlier in the year, but have only a modest effect on peak growing season length and little effect on annual growth.
This contrasts with the expectation *in most models* that an earlier start to growing season would significantly increase the rate of biomass accumulation and thereby enhance the CO~2~ sink of temperate deciduous forests.
It is, however, consistent with most tree-ring studies in temperate deciduous forests showing that annual growth tends to be more sensitive to climatic conditions during the peak growing season than in the spring[@helcoski_growing_2019], with eddy-covariance records showing that annual CO~2~ sequestration is more strongly linked to conditions in the peak growing season than to growing season length[@fu_maximum_2019], and with evidence that forest CO~2~ sequestration is sink-limited[@zani_increased_2020].
Thus, our findings indicate that warming spring conditions are unlikely to impact temperate deciduous forest growth sufficiently to produce any substantial negative climate change feedback.

@@

# (Introduction)

(*this will need to be shortened signficantly*)

@@

Reducing this uncertainty requires better understanding of the drivers of tree growth [@bugmann_tree_2019]. Tree growth is sensitive to climatic factors like precipitation, temperature, and cloudiness in early growing season [@helcoski_growing_2019]; [@Ciais_europe-wide_2005]; [@Goulden_exchange_1996], as well as other physical factors including nitrogen content of soil, increasing atmospheric CO~2~, deposition of atmospheric pollutants in acid rain [@mathias_disentangling_2018], and invasion by pests or pathogens. Understanding how growth is influenced by these factors is essential for developing accurate models to inform forest managers and policy makers about the role of future forests in climate change mitigation efforts.

## (spring temperatures are rising, and this is affecting leaf phenology and carbon cycling)
In recent decades, an increase in average spring temperature has been impacting many natural systems including forests [@parmesan_globally_2003]; [@zhizong_fingerprinting_2006]. 
The effects of this spring warming are most apparent when examining leaf phenology shifts, *which have been extensively studied ([@way_tree_2011]; [@klosterman_later_2018]), leading to many critical observations*. For example, some regions have seen earlier leaf out in deciduous trees, which is often matched or exceeded by delays in leaf senescence timing, when springtime temperatures are above average [@ibanez_forecasting_2010];@jeong_phenology_2011];@crabbe_extreme_2016]. The emergence of leaves signals the beginning of carbon sequestration or in other words, the growing season of a tree. It is believed that the longer the tree has leaves - or as the growing season is extended - the net annual carbon sequestered will increase as well (*i,e*, net ecosystem exchange, $NEE$; @churkina_spatial_2005; @keenan_net_2014; @zhou_explaining_2016].)

However, more of the interannual variation in forest $NEE$ is explained by variation in maximum C uptake rates than by the length of the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019]. This implies that earlier leafout will lead to smaller increases in NEE than changes in other phenological traits like the maximum C uptake rate. This idea agrees with patterns seen in tree cores, where peak growing season climate was more influential than spring climate [@helcoski_growing_2019] but many models still predict the extension of growing season will increase NEE. A paper by @zani_increased_2020 found that earlier leaf out led to earlier leaf senescence within the deciduous trees in their study, reducing the ability of the trees to sequester significantly higher levels of C. Instead they posit that senescence is induced when seasonal carbon-sink capacity of the tree is reached, no matter how extended the potential growing season is. This result directly contradicts those previous predictions that warmer Spring temperatures and earlier leaf out will lead to drastically increased C sequestration.

@@

# (Results)

## (Phenological patterns)

(*most of this is too detailed for Nature*)

## (Critical temperature windows)

## (Climate sensitivity of phenology )

## (Climate sensitivity of growth)

# (Discussion)

(*length may be okay, shouldn't be lengthened*)

Indeed, the tree-ring records reveal that *annual growth is most commonly controlled by climate conditions during the peak growing season*  [Fig. `r tree_rings`; @helcoski_growing_2019], .
(*1-2 sentences reviewing our regional analysis results (HOPEFULLY) generalizing about climate sensitivity of temperate deciduous forest growth.*)
Thus, although our dendrometer band analysis is limited to two sites, the finding that spring temperatures have limited effect on annual tree growth holds generally.

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (fig., REFS), an earlier start to net C sequestration (REFS), and--as shown here--earlier initiation of stem growth (Figs. `r climwin`- `r DOYtiming`).
Yet, an earlier start to the growing season does not imply a longer growing season [Table `r hypothesis_table`, Figs. `r pheno_Tsensitivity`, `r DOYtiming`;  @zani_increased_2020].
Rather, growth finishes sooner-- presumably because of *C sink limitation* [@zani_increased_2020].

# Methods

@@
